---
title: "Homework 3"
author: "Michelle Lee"
output: github_document
---
This is my solution to homework 3.
```{r import data}
library(tidyverse)
library(ggridges)
library(dplyr)
library(ggplot2)
library(p8105.datasets)
data("instacart")
names(instacart)
summary(unique(instacart$department))
summary(unique(instacart$aisle))
```


The data, instacart, shows the order details made on instacart. There are 1384617 rows and 15 variables and some of the key variables are, order_id, product_id, add_to_cart_order,reordered, user_id, eval_set, order_number, order_hour_of_day, days_since_prior_order, product_name, department_id, department. Based on the first set of data, we can infer that order id 1 ordered 8 total items, on the 10th hour of the day, 9 days after its previous order, ordered various items with corresponding aisles, which is in the training set of the evaluation. 

- How many aisles are there, and which aisles are the most items ordered from?
```{r}
#number of aisles;
ins_aisle= 
instacart %>%
  count(aisle)
  
ins_aisle
#finding the mode;
ins_aisle[order(ins_aisle$n,decreasing=TRUE),]
```
There are 134 aisles and fresh vegetables (n=150609) is the aisle where the most itemswere ordered from.

- Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
```{r}
ins_aisle2 =
  instacart %>% 
  group_by (aisle, department) %>% 
  summarize(number_products = n()) %>%
  arrange(department)

  ins_aisle2$aisle =
    factor(ins_aisle2$aisle, levels = unique(ins_aisle2 $aisle))
  
  ins_aisle2  %>% 
  ggplot(aes(x = aisle, y = number_products)) + 
  geom_point(aes(color = department)) +
  labs(title = "Products ordered per aisle within departments",
       y = "Products ordered", x = "Aisles") +
  scale_x_discrete(breaks = NULL) +
  theme(plot.title = element_text(hjust = 0.5))
```
- Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
```{r}
#create a subset;
popular_items <-
 instacart[which(instacart$aisle == "baking ingredients" |
                instacart$aisle == "dog food care" |
                instacart$aisle == "packaged vegetables fruits"),]

#drop other columns and find frequency of each product;
popular_items=
  select(popular_items, aisle, product_name) %>%
  count(aisle,product_name) 

#sort in decreasing order;
popular_items= 
  popular_items[order(popular_items$n,decreasing=TRUE),]

#create a table;
require(data.table)
popular_items_top3 =
  data.table(popular_items, key = "aisle")

popular_items_top3=
  popular_items_top3[,head(.SD,3), by=aisle]
knitr::kable(popular_items_top3)

```

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).
```{r pla_cic}
pla_cic <-
 instacart[which(instacart$product_name == "Pink Lady Apples" |
                instacart$product_name == "Coffee Ice Cream"),]
pla_cic

#drop other columns;
pla_cic =
select(pla_cic , order_hour_of_day, order_dow, product_name)
pla_cic

#group by each dow;
require(data.table)
by_dow = 
  pla_cic %>%
  group_by (order_dow, product_name)
by_dow =
  by_dow %>% summarise(
  mean_hours = round(mean(order_hour_of_day))
  )
knitr::kable(by_dow)
```
#problem 2
Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).
```{r import data for acc}
#import csv
acc_data = read_csv("./accel_data.csv")
acc_data_clean = janitor::clean_names(acc_data)
acc_data_clean
acc_data_clean =
  pivot_longer(
    acc_data_clean,
    activity_1:activity_1440,
    names_to = "minute",
    values_to= "activity_counts")

acc_data_clean =
  acc_data_clean %>%
  mutate(weekday_weekend = day) %>%
  mutate(weekday_weekend = recode(weekday_weekend, 
                                  'Monday' = 'weekday',
                                  'Tuesday' = 'weekday',
                                  'Wednesday' = 'weekday',
                                  'Thursday' = 'weekday',
                                  'Friday' = 'weekday',
                                  'Saturday' = 'weekdend',
                                  'Sunday' = 'weekend'
                                  )) 
acc_data_clean$day <-
  factor(acc_data_clean$day, levels= c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday","Sunday"))

acc_data_clean[order(acc_data_clean$day), ]
view(acc_data_clean)
summary(acc_data_clean)
```
Data description: 
This dataset contains information regarding five weeks of accelerometer data collected on a 63 year-old male with BMI 25. (After pivoting,) The dataset contains a total of 50400 observations and 6 variables, including new variables, activity_counts, minute, and weekday vs weekend. There is no missing data. “activity_counts” is the activity count for a given mintue in the observation.

Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?
```{r}
total_activity =
acc_data_clean %>%
    group_by(week,day) %>%
    summarise(
      total_activity = sum(activity_counts)
    ) %>%
  pivot_wider(
    names_from = "day",
    values_from = "total_activity"
  ) 
knitr::kable(total_activity)
```
Based on the table above, it is difficult to identify any clear trends betwen total activity variable for each day.Generally, this person had a consistent total activity between 340115~ 468869 on Wednesdays. On Saturday of week 4 and 5, this person had low activity counts compared to other days.

Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.
```{r making a single panel plot}
acc_data_clean %>%
    ggplot(aes(x = minute, y = activity_counts,group=day)) + 
  geom_line(aes(color = day)) +
  labs(title = "24 hour activity time courses over 5 weeks", 
       x = "Hours", 
       y = "Activity Counts") 
```
#problem 3
```{r import data and write a short description}
library(p8105.datasets)
data("ny_noaa")
ny_noaa= 
  ny_noaa[order(ny_noaa$date,decreasing=FALSE),]
ny_noaa[order(ny_noaa$date,decreasing=TRUE),]
ny_noaa
unique(ny_noaa$id)
summary(is.na(ny_noaa))
```
Description of the dataset:
There is a total of 2,595,176 total observations(rows) and 7 variables (columns), which includes ID, date, prcp, snow, snwd, tmax, and tmin. The important variables are date, prcp, snow, tmax, and tmin, as these datapoints indicate the weather on a certain day. The date ranges from 1981-01-01 to 2010-12-31, but not all days between that period are recorded. There are 746 unique IDs/ locations. Variables prcp (145838), snow (381221), snwd (591786), tmax(1134358), and tmin(1134420) are missing values.

Do some data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. 
``` {r data cleaning}
ny_noaa_clean =
  ny_noaa %>%
  mutate(
    month = lubridate::floor_date(date, unit = "month"),
    year = lubridate::floor_date(date, unit = "year"),
    day = lubridate::floor_date(date, unit = "day")
  ) %>% 
  separate(year, into = c("year", "trash1", "trash2"), sep = "-") %>%
  separate(month, into = c("trash3", "month", "trash4"), sep = "-") %>% 
  separate(day, into = c("trash5", "trash6", "day"), sep = "-") %>% 
  select(-contains("trash")) %>% 
  mutate(tmax = as.numeric(tmax), tmin = as.numeric(tmin)) %>% 
  mutate( prcp = prcp/10, tmax = tmax/10, tmin = tmin/10)
```
For snowfall, what are the most commonly observed values? Why?
```{r snowfall}
ny_noaa_snow_clean = 
  na.omit(ny_noaa_clean, cols ='snow')
#find frequency of each values in snow;
mode_snow =
  ny_noaa_snow_clean %>%
  count(snow)

#sort in decreasing order;
mode_snow= 
  mode_snow[order(mode_snow$n,decreasing=TRUE),]
```
The most commonly observed values are 0 (1112758), 25(15809), and 13 (12460) in the snow variable. The high observed value of 0 could be explained by the fact that many parts of NY do not snow during much of the year, except during the winter.  
Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?
```{r}
jan_july = 
  ny_noaa_clean[which(ny_noaa_clean$month == "01" |
                ny_noaa_clean$month == "07"),]
jan_july

#drop other columns and change variables to int;
jan_july=
  jan_july %>%
    select(id, year, month, tmax) %>%
    mutate(year = as.numeric(year)) %>%
    summarize(avg_tmax = mean(tmax, na.rm = TRUE)) %>% 
    ggplot(aes(x = year, y = avg_tmax)) +
    geom_point(aes(color = id)) +
    geom_smooth()+
    facet_grid(.~month) +
    labs( title = "Temperature in January and July by year", y = "Average Maximum Temperature", x = "Time (year)") +
    theme(legend.position = "none", panel.spacing = unit(2, "lines"), plot.title = element_text(hjust = 0.5))
```
There is a general trend for the average maximum temperature in January and July in each station across years.For January of each year, there has been a cyclical fluctuation in maximum temperature over the years. Based on the graph, we can infer that there seems to be a peak in mean max temp every 10 years, and there are a few outliers for the trend in January. For July, the average max temperature has remained between 25-30º C,  but there is a notable low average maximum temperature in July that occured between 1985 and 1990.

Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.
```{r }
tmax_tmin=
  ny_noaa_clean %>%
    select(tmax,tmin) %>%
    na.omit(ny_noaa_clean, cols ='tmax','tmin') %>%
    ggplot(aes(x = tmax, y = tmin)) + 
  geom_hex() +  
  labs(title = "tmax vs. tmin", x = "tmin (ºC)", y = "tmax (ºC)") +
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5))
tmax_tmin
```
Based on 1,458,900 observations of tmax and tmin, excluding NAs. 

```{r}
snow_year = 
  ny_noaa_clean %>% 
  filter(0 < snow & snow < 100) %>%
  ggplot(aes( x = year, y = snow)) +
  geom_hex() + 
  scale_x_discrete(breaks = c("1981","1990","2000","2010")) +
  labs(title = "Snowfall by year", y = "snow fall (mm)")+
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5))
snow_year

```
